Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-08-11 JHEP_submission/emu.tex
minor changes
minor diction

\documentclass[11pt,letter]{article}
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scaleBox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scaleBox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

@@

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  
We exploit special properties of all even-point theories towards establishing an efficient unitarity based amplitude construction. 
In doing so, we find evidence that the leading IR divergence of NLSM amplitudes exponentiates when the target space is $\mathbb{CP}^1\cong SU(2)/U(1)$. We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms needed to restore $U(1)$ invariant behavior at loop-level can be constructed via a symmetric-structure double-copy.  
We also demonstrate that the divergent part of the one-minus $(-+++)$ two-loop anomaly vanishes upon introducing an evanescent operator. In addition to these pure photon counterterms, we verify through explicit calculation that the anomalous matrix elements that violate $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. 
Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes permit double-copy construction through two-loop order by reproducing our unitarity based result with a double copy between color-dual $\mathcal{N}=4$ super-Yang-Mills  and our two-loop NLSM amplitudes. 
This result supports the possibility of  color-dual representations for NLSM beyond one-loop. 
We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

\begin{document}
\maketitle
\flushbottom

@@

For example, consider the two-loop divergence for the $(++++)$ anomalous matrix element in pure BI theory. The 4D helicity structure can be captured by a symmetric structure double-copy between NLSM pion numerators and a local gauge theory contact at $\mathcal{O}(\alpha'^5)$. The $s$-channel numerators for this symmetric-structure double copy are as follows
\begin{equation}\label{eq:symNum2loop}
n_s^{\text{dd,}\pi} = s^2 \qquad n_s^{\text{dd,HD}} = s^5 \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
Double-copying these kinematic numerators yields a matrix element of the form:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{BI}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{n_g^{\text{dd,HD}}n^{\text{dd,}\pi}_g}{d_g} = (s^6+t^6+u^6) \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
While this construction lacks any algebraic relations between the four-point kinematic-factors, similar symmetric numerators were found at six-point for NLSM, which obey non-trivial algebraic relations \cite{Carrasco:2022jxn}. In addition to the photon counterterms constructed above, symmetric double copy is likewise a natural description of gravitational counterterms. Rather than composing the local vector numerators, $n^{\text{dd,HD}}$, with symmetric NLSM numerators, $n^{\text{dd,}\pi}_g$, they can also be squared -- yielding a gravitational counterterm:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{GR}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{(n^{\text{dd,HD}}_g)^2}{d_g} = (s^9+t^9+u^9)  \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
In general, any four-point symmetric vector numerator can constructed from two $D$-dimensional gauge theory building blocks \cite{Carrasco:2022jxn},
\begin{equation}
n_s^{\text{dd,vec,1}} = f_{12}f_{34}\,, \qquad n_s^{\text{dd,vec,2}} = f_{1324}\,,
\end{equation}
by composing with the symmetric scalar numerators, $n_s^{\text{dd,1}} = s$ and  $n_s^{\text{dd,2}} = tu$. With these building blocks, symmetric double copy construction captures the exceptional four graviton amplitude of Ref.~\cite{Boels:2016xhc}, which only considered linear combinations of gauge theory amplitudes rather than gauge theory numerators. We note that a similar construction was recently identified in Ref.~\cite{Bourjaily:2023tcc} beyond four-point, which demonstrated that gravity amplitudes permit double copy construction in terms of gauge-invariant numerators that do not obey any algebraic relations between themselves, much like the symmetric numerators above in \eqn{eq:symNum2loop}. We see further exploring higher multiplicity examples of symmetric double copy as an important direction of future study.

>>>>

\documentclass[11pt,preprint]{article}
\pdfoutput=1
\usepackage{jheppub}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{mathrsfs}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{tocloft}
\renewcommand{\cftdot}{.}
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scaleBox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scaleBox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

@@

\abstract{
We identify novel  structure in newly computed multi-loop amplitudes and quantum actions for even-point effective field theories, including both the nonlinear sigma model (NLSM) and double-copy gauge theories such as Born-Infeld and its supersymmetric generalizations.  
We exploit special properties of all even-point theories towards establishing an efficient unitarity based amplitude construction. 
In doing so, we find evidence that the leading IR divergence of NLSM amplitudes exponentiates when the target space is $\mathbb{CP}^1\cong SU(2)/U(1)$. We then systematically compute the two-loop anomalous behavior of Born-Infeld, and find that the counterterms needed to restore $U(1)$ invariant behavior at loop-level can be constructed via a symmetric-structure double-copy.  
We also demonstrate that the divergent part of the one-minus $(-+++)$ two-loop anomaly vanishes upon introducing an evanescent operator. In addition to these purely photonic counterterms, we verify through explicit calculation that the anomalous matrix elements that violate $U(1)$ duality invariance can be alternatively cancelled by summing over internal $\mathcal{N}=4$ DBIVA superfields. 
Finally we find that $\mathcal{N}=4$ Dirac-Born-Infeld-Volkov-Akulov (DBIVA) amplitudes admit double-copy construction through two-loop order by reproducing our unitarity based result with a double copy between color-dual $\mathcal{N}=4$ super-Yang-Mills  and our two-loop NLSM amplitudes. 
This result supports the possibility of  color-dual representations for NLSM beyond one-loop. 
We conclude with an overview of how $D$-dimensional four-photon counterterms can be constructed in generality with the symmetric-structure double-copy, and outline a convenient way of counting evanescent operators using Hilbert series as generating functions. }

@@

\begin{document}
\maketitle

@@

For example, consider the two-loop divergence for the $(++++)$ anomalous matrix element in pure BI theory. The 4D helicity structure can be captured by a symmetric structure double-copy between NLSM pion numerators and a local gauge theory contact at $\mathcal{O}(\alpha'^5)$. The $s$-channel numerators for this symmetric-structure double copy are as follows
\begin{equation}\label{eq:symNum2loop}
n_s^{\text{dd,}\pi} = s^2 \qquad n_s^{\text{dd,HD}} = s^5 \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
Double-copying these kinematic numerators yields a matrix element of the form:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{BI}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{n_g^{\text{dd,HD}}n^{\text{dd,}\pi}_g}{d_g} = (s^6+t^6+u^6) \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
While this construction lacks any algebraic relations between the four-point kinematic-factors, similar symmetric numerators were found at six-point for NLSM, which obey non-trivial algebraic relations \cite{Carrasco:2022jxn}. In addition to the photon counterterms constructed above, symmetric double copy is likewise a natural description of gravitational counterterms. Rather than composing the local vector numerators, $n^{\text{dd,HD}}$, with symmetric NLSM numerators, $n^{\text{dd,}\pi}_g$, they can also be squared -- yielding a gravitational counterterm:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{GR}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{(n^{\text{dd,HD}}_g)^2}{d_g} = (s^9+t^9+u^9)  \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
In general, any four-point symmetric vector numerator \cite{Carrasco:2022jxn} can be constructed from two $D$-dimensional gauge theory building blocks,
\begin{equation}
n_s^{\text{dd,vec,1}} = f_{12}f_{34}\,, \qquad n_s^{\text{dd,vec,2}} = f_{1324}\,,
\end{equation}
by composing with the symmetric scalar numerators, $n_s^{\text{dd,1}} = s$ and  $n_s^{\text{dd,2}} = tu$. For example, with these building blocks, symmetric double-copy construction captures the exceptional four graviton amplitude of Ref.~\cite{Boels:2016xhc}, which only considered linear combinations of gauge theory amplitudes rather than gauge theory numerators. We note that a similar construction was recently identified in Ref.~\cite{Bourjaily:2023tcc} beyond four-point, which demonstrated that gravity amplitudes permit double copy construction in terms of gauge-invariant numerators that do not obey any algebraic relations between themselves, much like the symmetric numerators above in \eqn{eq:symNum2loop}. We see further exploring higher multiplicity examples of symmetric double copy as an important direction of future study.
